---
title: "LSTAT2120 - Linear Project"
author: "Lamy Lionel, Aurèle Bartolomeo, Rémi Gengler"
date: "26/12/2020"
output:
  prettydoc::html_pretty:
    css : "sources/main.css"
    theme: cayman	
    highlight: github	
    toc: true
    toc_depth: 3
    fig_caption: true	
    number_sections: no
  html_document: default
---

```{r global.options, include=F}
knitr::opts_chunk$set(
  cache = TRUE,
  echo = F,
  eval=T,	
  	
  warnings=F,	
  message=F,	
  	
  out.width= "80%",	
  fig.align = "center",	
  fig.path = "figs/",	
  fig.keep = "all"	
  	
)	
rmarkdown::html_document(df_print = "formattable")
```

```{r options}
opt.digits = 3
opt.barcolor = "#0a4d7d"
```

```{r utils}
models.comparison = data.frame()

addComparison = function(model, model_name){
  to_combine = data.frame(
      LogLik = ifelse(is.null(logLik(model)[1]), yes=NA, no=round(logLik(model)[1], opt.digits)),
      AIC = round(AIC(model), opt.digits),
      #BIC = round(BIC(model), opt.digits),
      R2 = round(summary(model)$r.squared, opt.digits),
      Adjusted.R2 = round(summary(model)$adj.r.squared, opt.digits),
      Deviance = ifelse(is.null(deviance(model)), yes=NA, no=round(deviance(model), opt.digits)),
      #Deviance.custom = cost_deviance(base_train_freq$Frequency, model$fitted),
      row.names = as.character(model_name)
  )
  rbind(
    models.comparison,
    to_combine
  )
}
```

# Intro and objective

For this project, we will use a dataset that contains all the FIFA 19 player characteristics (FIFA is a football simulation video game, and our dataset come from the 2019's edition). This dataset contains 18059 observations (all the different players created in the game) and 34 variables (characteristics such as Age, Nationality, Wage, Stamina, ...). 

The purpose of our project is to predict the wage of a player based on all his features. This is a very interesting question because it could help club managers to know if some of their players are being underpaid (that would push the player to leave) or overpaid (that could be a threat to club finances). It could also help recruiters to have an idea of the price to pay for a player, depending on the features he is looking for. Finally, it would be a helpful tool for young players, for example to know on which feature they have to work in order to have their wage increased.

For this purpose we will use linear models only. So for a start we will look more precisely at our dataset and see if the classical assumptions for linear models are respected. We will also check for nonlinearity, influential observations, multicollinearity, heteroskedasticity and autocorrelation. All these concepts will be reexplained in the appropriate section. If some hypotheses seem clearly not satisfied, we will then take remedial actions. We will retain different models based on several regression methods seen in course and will compare them based on robust criterions (for example, their ability to predict correctly th value of the wage for a completely new player).

# Data cleaning and separtion of the dataset

We start by doing a complete case analysis (i.e eliminating all observations that contain missing values). We can do that without intoducing any bias in our analysis, because observations containing missing values constitute an extremely small portion of our dataset (48 of 18207 observations, so less than 0.3%). We than improve the usability of the dataset by changing some units. Initially, the variable Nationality (that recorded the nationality of the player) was encoded as the country of the player, but there were then too much levels and some of them seemed to be irrelevant, so we decided to modify it manually and replace these levels by the associated continents.
Finally, we separate randomly 20% of our observations. These observations will be used for prediction, but not for model estimation. 

```{r data_setup}
set.seed(28122020)
original = read.csv("./sources/foot_last.csv", sep=";", encoding = "utf8", skipNul = T, stringsAsFactors = F)
original = na.omit(original)
original = original[-which(original == 0, arr.ind = T)[,1],] # remove rows where value is 0

# Dummies variables
dummies = data.frame(model.matrix(~ Nationality - 1, data=original), 
                     model.matrix(~ Preferred_Foot - 1, data=original))

# Transformations
dataset = within(original, {
  Weight  = round(as.numeric(Weight*0.453592))
  Height = round(sapply(Height, function(x) 
    as.numeric(strsplit(x, "'")[[1]][1])*12 + as.numeric(strsplit(x, "'")[[1]][2])*2.54))
  
  Value = Value_M #* 10^6
  Wage = Wage_K #* 1000
  
  Value_M = NULL
  Wage_K = NULL
  Nationality = NULL
  Preferred_Foot = NULL
})

```

# Descriptive analysis

Among the 34 variables, 27 take values in percentage : it mainly concerns the game features such as agility, dribbling, ball control... The other quantitative variables are the age (given in years), the wage (given in Euros), the height (given in centimeters), the weight (given in kilograms) and the value of the player (given in Euros). We should also emphasize the fact that we have two categorical variables called Preferred Foot (with two levels : Left and Right) and Nationality (with five levels : Africa, America, Asia, Europe and Oceania). We can look at the two corrisponding barplot. We see that more than half of the players are europeean, while the 4 other continents don't exceed 25% individually. We also observe that players that are right-footed are almost four times more than left-footed players.

```{r descriptive, out.width="80%"}
# Histogram of qualitative variable

par(mfrow=c(1,2))
barplot(prop.table(table(original$Nationality)), main="Nationality of the player", col = opt.barcolor, ylab="Percentage",las=2)
barplot(prop.table(table(original$Preferred_Foot)), main="Preferred foot of the player", col = opt.barcolor, ylab="Percentage")

# Mean, Sd, Skewness, Kurtosis
library(moments)
library(formattable) # Better print
library(ggplot2)
library(gridExtra)

desc.quantitative = (function(data){
  
  addStats = function(col, name){
    
    return(data.frame(
      Variable=name,
      Mean = round(mean(col, na.rm = T), opt.digits),
      Std.Deviation = round(sd(col), opt.digits),
      Skewness = round(skewness(col), opt.digits),
      Kurtosis = round(kurtosis(col), opt.digits)
      ))
  }
  
  for (c in colnames(data)){
    if (exists("results")){
      results = rbind(results, addStats(data[[c]], c))
    }else{
      results = addStats(data[[c]], c)
    }
  }
  return (results)
   
})(dataset)

desc.quantitative.order = order(as.character(desc.quantitative$Variable))
desc.quantitative = desc.quantitative[desc.quantitative.order,]
row.names(desc.quantitative) = NULL # hide row number

formattable(desc.quantitative,
            align=c("l", rep("c", ncol(desc.quantitative)-1))
            #,list(Skewness = formatter("span", style = x ~ style(color= ifelse(abs(x) > 0.5, ifelse(abs(x) >= 2,"red", "orange"), "black"))))
            )
```


```{r descriptive_graph, out.width="100%", fig.height=1.8}
# Boxplot and Histogram

desc.quantitative.boxplots = function(data, nrow=2, ncol=2, boxframe=T){
  
  
  layout(mat=matrix(seq(nrow*ncol*2), 2*nrow, 1*ncol, byrow=F), heights = rep(c(4,1.5), ncol))
  drawPlots = function(coname){
    
    par(mar=c(0, 2, 2, 2))
    hist(data[[coname]], axes=T, main=coname, col = opt.barcolor, xlab="", ylab="Count", xaxt="n")
    par(mar=c(3, 2.3, 0, 2.3))
    boxplot(data[[coname]], horizontal = T, frame=boxframe)
    
  }
  
  invisible(sapply(colnames(data), function (x) drawPlots(x)))
  
}

desc.quantitative.boxplots(data.frame(dataset, Wage.log=log(dataset$Wage), Value.log=log(dataset$Value)),
                           1, 4, T)
  
```

Now we can take a look at the table that shows some interesting descriptive values about our quantitative variables. More precisely, we have calculated the mean, the standard deviation, the skewness and the kurtosis for each of them. It would be too massive to describe precisely each result of this Table since we have a lot of variables, but we will try to outline the most important ones. 

Let's start with the Age variable. Players age have mean 25, with a 4.7 standard deviation. We see that the variable is skewed right (nothing surprising, some players can still play at 40 but none can play at 5). The kurtosis is a bit negative (recall that in SAS, a normal distribution has kurtosis 0, not 3), that means that the variable is maybe a bit light-tailed. 

Now we can look at our variable of interest, the wage. It seems that it is not at all following a normal distribution. In fact, we have a a skewness of 7.8 which means that it is very right-tailed. Furthermore, it has a kurtosis of 103 so the wage variable is extremely thick in the tails. We can back up this claim by looking at the histogram and boxplot of the wage. This variable is definitely not normal.

Similarly from the wage variable, we see that the value of the player is also right-skewed and very thick in its tails. Again, we can look at the histogral/bowplot figure of the value variable and we confirm this tendency.

Weight and height of the players seem to be normally distributed (maybe a bit right-skewed for the weight) around 181 cm and 75 kg, respectively.

The remaining variables are all about players abilities. Like we have already said, these variables take values between 0 and 100. We will not describe each of them specifically, since they are all quite similar. In general, they have mean between 45 and 65 and standard deviation between 10 and 20. In addition, they have skewness near 0 and kurtosis near 3. 

We now look at the correlation matrix. The aim of this step is to detect high pairwise correlations (and therefore multicollinearity). Note that even if there were no pairwise correlations, it would not mean that there is no multicollinearity. In fact, there can be multicollinearity invisible in the correlation matrix because multicollinearity can come frome linear dependance between 3 or more variables.

```{r full_corrplot}
library(corrplot)
corrplot(cor(dataset), type="upper", order="hclust", tl.col="black", tl.srt=90, tl.cex = 0.7)
```

We have here the upper part of the correlation matrix. We see that all variables in the bottom right have very high pairwise correlations (lots of those correlations are even superior to 0.8). We are not very surprised of that, because we could expect that good players have globally good abilities and conversely, bad players have on average bad abilities. Thus, those variables contains almost the same information as what we already have in the overall variable. Considering this fact, we decided to discard those variables in order to avoid too much multicollinearity problems. 

After this suppression, we see that the following correlation matrix is better, in the sense that we have no longer blocks of high correlated variables. 

```{r min_corrplot}
# Suppression/merge of correlated variables
clean_dataset = dataset[,c(1:5, 8, 19, 22, 24, 26:27, 31, 32:34)]
corrplot(cor(clean_dataset), type="upper", order="hclust", tl.col="black", tl.srt=90, tl.cex = 0.7)
```

Again, this does not mean that we have no multicollinearity at all. We will look at that before model selction in the next section, with more sophisticated tools. 


# Model Selection

We will now start to select variables to include in our model. We will use the two following types of model selction : a stepwise selection using AIC and a LASSO procedure. 

```{r}
library(olsrr)
library(lmtest)

library(MASS)
library(glmnet)

# Splitting into training and validation
tmp = sample(nrow(clean_dataset), round(.80*nrow(clean_dataset)), replace=F)

train = cbind(clean_dataset[tmp,], dummies[tmp,])
valid = cbind(clean_dataset[-tmp,], dummies[-tmp,])

rm(tmp)

formula.complete = as.formula("Wage ~ .")
formula.complete.logy = as.formula("log(Wage) ~ .")

formula.without = as.formula("Wage ~ . - Value")

model.complete = lm(formula.complete, data=train)
model.complete.logy = lm(formula.complete.logy, data=train)

model.without = lm(formula.without, data=train)

models.comparison = addComparison(model.complete, "Complete")
models.comparison = addComparison(model.complete.logy, "Complete log(Y)")
```



## Stepwise Regression using AIC

We will use a stepwise function that finds a model that minimizes AIC, using a step-by-step approach. As a reminder, the AIC, just like the adjusted R-squared, is a measure of the quality of the model that penalizes the addition of new variables. This is what we need in order to do a stepwise selection. The method will check wheter the AIC keeps decreasing while we include more and more variables. 


```{r model_selection stepwise}
# Stepwise AIC 
aic.mass = stepAIC(model.complete, direction = "both", trace = F)
aic.mass

formula.aic = as.formula("Wage ~ Age + Overall + Potential + Weight + HeadingAccuracy + 
    Reactions + Strength + Interceptions + Value + NationalityAmerica")


models.comparison = addComparison(lm(formula.aic, data = train), "Stepwise AIC")

# Stepwise AIC log(Y)
aic.masslogy = stepAIC(model.complete.logy, direction = "both", trace = F)
aic.masslogy

formula.aiclogy = as.formula("log(Wage) ~ Overall + Potential + Weight + HeadingAccuracy + 
    Reactions + Jumping + Strength + Aggression + Composure + 
    Marking + Value + NationalityAfrica + NationalityAsia + NationalityEurope + 
    Preferred_FootLeft")


models.comparison = addComparison(lm(formula.aiclogy, data = train), "Stepwise AIC log(Y)")
```

In the stepwise summary above we see that the selection discards 6 variables : Age, Height, Weight, Jumping

## LASSO Procedure

The LASSO estimator is comprable to the OLS in the sense that the goal is to minimize the sum of squared residuals. But the main difference lies in the fact that it imposes a constraint on the L1 norm of the model parameter $\beta$.
Indeed, for a certain coefficient t>0 to be determined, we impose that \[ \sum_{j=1}^{p} |\beta_j| \leq t \].


```{r}
#lasso
lasso.cv = cv.glmnet(x=data.matrix(within(train, {Wage = NULL})), y = train$Wage, family = "gaussian", alpha = 1)
lasso.coef = coef(lasso.cv, s="lambda.min")

formula.lasso = as.formula("Wage ~ Age + Overall + Potential + Height + Weight + HeadingAccuracy + Reactions + Jumping + Strength + Interceptions + Composure + Marking + Value + NationalityAfrica + NationalityAmerica + NationalityEurope + NationalityOceania + Preferred_FootLeft")
model.lasso = lm(formula.lasso, data=train)

models.comparison = addComparison(model.lasso, "Lasso")



#lasso log(Y)
lasso.cvlogy = cv.glmnet(x=data.matrix(within(train, {Wage = NULL})), y = log(train$Wage), family = "gaussian", alpha = 1)
lasso.coeflogy = coef(lasso.cvlogy, s="lambda.min")

formula.lassology = as.formula("log(Wage) ~ Age + Overall + Potential + Height + Weight + HeadingAccuracy + Reactions + Jumping + Strength + Aggression + Composure + Marking + Value + NationalityAfrica + NationalityAmerica +  NationalityAsia + NationalityEurope + NationalityOceania + Preferred_FootLeft + Preferred_FootRight")
model.lassology = lm(formula.lassology, data=train)

models.comparison = addComparison(model.lassology, "Lasso log(Y)")
```

## Final Model (without interactions)
We have now looked at 6 differents models. The complete regression model, the AIC-stepwise regression model and the lasso regression model. For each of them, we have estimated a model of regression against tha wage variable and against the log_wage variable. We have done that because we suspected non linear relations in our model. 
```{r}
models.comparison
```

The table above show a model comparison for those 6 models. We use various criteria such as the loglikelihhod, the AIC, the adjusted R-squared and the deviance. We see that the R-squared is slightly better if we don't take the log of the response variable, but all the other criteria are far much better when we take it. We decide to take as final model the AIC-stepwise model because it has almost as good goodness of fit criteria as the LASSO, but it has only 15 variables (compared to 20 for the LASSO regression).

```{r}
#setting the final model (aic logy)
model = lm(log(Wage) ~ Overall + Potential + Weight + HeadingAccuracy + 
    Reactions + Jumping + Strength + Aggression + Composure + 
    Marking + Value + NationalityAfrica + NationalityAsia + NationalityEurope + 
    Preferred_FootLeft, data = train)

```

## Interactions

We will test the interaction between the preferred foot and the marking variable.

```{r}
model.interaction = lm(log(Wage) ~ Overall + Potential + Weight + HeadingAccuracy + 
    Reactions + Jumping + Strength + Aggression + Composure + 
    Marking + Value + NationalityAfrica + NationalityAsia + NationalityEurope + 
    Preferred_FootLeft + Preferred_FootLeft * Marking, data = train)

models.comparison = data.frame()
models.comparison = addComparison(model, "Model")
models.comparison = addComparison(model.interaction, "Model with interaction")

```

In the comparative table above, we see that the interaction we are testing does not improve our criteria of goodness of fit. Adjusted R-squared is the same as for the model without interaction, and the AIC is even slightly greater. Therefore we will working with the model without interaction with the PreferredFoot.

# Underlying Hypotheses Testing

We will now check for nonlinearity, outliers (and influential observations), multicollinearity, heteroskedasticity, autocorrelation and normality of the residuals. If some of these hypotheses are not fullfilled, we will try to take remedial actions.

## Nonlinearity

We would like to chek wheter it was a good choice to make a linear model regression. In particular, we would like to check if the regression function is linear. In order to do that, we will look at scatter plots of the residuals $e_i$ against $X_{ij}$, with j = 1, ..., p-1.  

```{r}
#train$Value = log(train$Value)
par(mfrow=c(4,4))
for(i in 1:15){
  par(mar=c(2,2,2,1))
  plot(x = train[,i], y=lm(as.formula(paste0("log(Wage) ~ ", colnames(train)[i])), data=train)$residuals, main = colnames(train)[i], ylab = "Residuals", pch=16, col=opt.barcolor)
}

```

We are looking for non-linear patterns on the above plots of the residuals and, if necessary, take remadial actions to cope with that problem. For example, if we see a quadratic scatterplot, we would be tempted to add a quadratic term for this exmplaining variable. In our scatterplots, we don't see evidence of hidden qudratic relations, altough 
the scatterplots for the Overall and Value variables tends to show that there is probably a non-linear relation between those terms and our variable of interest. We choose to keep going with the intial model, because the non-linear formula seems to be not straightforward, and therefore we fear to complicate too much the model if we try to deal with it. In addition, recall that we have already taken the log of the wage variable, so we have not so much possibilities remaining.

## Outliers

We will first check for outliers with respect to X. For that, we calculate the levrages that we can find with the following formula 
\[h_{ii}=(1, X_{i1}, ..., X_{i,p-1})(X'X)^{-1}(1, X_{i1}, ..., X_{i,p-1})'\].
In fact, they are the elements on the diagonal of the rojection matrix $H=X(X'X)^{-1}X'$. Then, we say that the observation $X_i$ is an outlier if $h_{ii} > \frac{2p}{n}$.


Then, we check for outliers with respect to Y (the wage). We say that $Y_i$ is an outlier if 
\[|d_i^*| > t_{n-p-1; 1-\alpha/2}\], where \[d_i^*=e_i \Bigg(\frac{n-p-1}{SSE(1-h_{ii})-e_i^2} \Bigg)^{1/2}\].


We now look at the influential observations for the fitted values. Recall that the i-th observation is influential if $\big|DFFITS_i\big|>2\sqrt{p/n}$, where $DFFITS_i=d_i^*\sqrt{\frac{h_{ii}}{1-h_{ii}}}$. 


In addition to that kind of influential observations, we have to consider the influential observations for the regression coefficients. The i-th observation is influential if $\big|DFBETAS_{k,i}\big|>2/\sqrt{n}$, where $DFBETAS_{k,i}=\frac{\hat{\beta}_k-\hat{\beta}_{k(i)}}{\sqrt{MSE_{(i)}c_k}}$ and $c_k=(X'X)_{kk}^{-1}$.

# Multicollinearity

We will use the Variance Inflation Factors (VIF) to determine if there is multicollinearity problems. Recall that the VIF for $\hat{\beta}_k$ is defined by $VIF_k = \frac{1}{1-R_k^2}$ for $k=1,..., p-1$, where $R_k^2$ is the coefficient of determination of a regression of $X_k$ on $X_1,...,X_{k-1},X_{k+1},...,X_{p-1}$. The decision rule to decide if there is a multicollinearity problem is the following : if either the maximum VIF is higher than 10 or if the average VIF is considerably larger than 1 (A $VIF_k$ greater than 10 means that the corrisponding $R^2_k$ is larger than 0.9, which is a sign of approximate collinearity). This method, instead of the pairwise correlation matrix that we analysed before, has the advantage to detect also multicollinearity relations involving more than 2 variables.

```{r}
ols_vif_tol(model)
```

We see on the result table above that we have no VIF superior to 10, which is good news. The mean of all these VIF is of 2.44. Since it is not extremely far from one, we decide not to do a Ridge Regression, indeed it seems not necessary so far.

# Heteroskedasticity 

We plot the residuals against the fitted values to consider possible heteroskedasticity. If the variance of the residuals is considerably variable, we discard the hypothesis of homoskedasticity, meaning that there is heteroskedasticity. We can also look at the Breusch-Pagan test to detect heteroskedasticity. Note that the null hypothesis of this test states the homoskedasticity whereas in the alternative hypothesis the variance of the residuals is dependent on the observation (i.e. player) considered (equivalent to heteroskedasticity).

```{r}
#Residuals vs fitted
plot(model$residuals, model$fitted.values)
```
It seems that the more the fitted values are big, the more our residuals tends to be negative. It is a bit complicate to decide on the sole basis of this graph wheter there is heteroscedasticity, so we decide to do a Breusch-Pagan test.

```{r}
ols_test_breusch_pagan(model, rhs = T, multiple=T)

```

With a p-value of $3.6e^{-174}$, it is perfectly clear that we can reject the null hypothesis of homoscdasticity. Hopefully, we have done a robust inference so our $\hat\beta$ estimator is consistent.


# Autocorrelation

Autocorrelation is discussed through the Breusch-Godfrey test. Recall that this test looks at non-autocorrelation ($H_0$) versus autocorrelation ($H_1$). The null hypothesis can be written as $Corr(\varepsilon_t, \varepsilon_{t-k})=0$ for $k=1,...,p$.

```{r}
bgtest(model)

```

The p-value of this Breush-Godfrey test tells us that we can not reject the null hypothesis of no autocorrelation. Therefore, we don't take remedial actions in this case.



# Normality of the residuals

In order to test the normality of the residuals, we can first consider a Quantile-Quantile plot of the residuals. In this plot, we compare the empirical quantiles of the residuals with those of a normal distribution. 

```{r}
qqnorm(model$residuals, pch = 1, frame = FALSE, col = opt.barcolor)
qqline(model$residuals, lwd = 2)

```

Therefore, the points on this graph should follow the straight red line. We see that this is not the case for the first quantiles, even though it seems to get better and better then. We will do a Jarque-Bera test to take the decision of rejecting normality of the residuals or not. This test compares the coefficients of skewness $S$ and kurtosis $\kappa$ with the theoreticla values for a normal distribution. Recall that for a normal distribution, skewness is equal to 0 and kurtosis is equal to 3 (this is our null hypothesis $H_0$). The Jarque-Bera test statistic is then given by 
\[JB = \frac{n}{6} \Big[\hat{S^2}+\frac{(\hat\kappa-3)^2}{4}\Big] \sim \chi^2_2\], under $H_0$.

```{r}

jarque.test(model$residuals)
```

With the result of the test above, we reject the normality of the error terms. This could be the result of an inadequate model, in fact this could mean that the error is not random. Like we have already said in the non-linearity section, a complicated non-linearity relation could maybe fix this problem, but it would be out of the scope of this project. 



# Significance of the Coefficients
We show below the estimates and significance test for the variables that are included in our model. 


```{r}
summary(model)
```

The variable weight, jumping and PreferredFoot seem to be not significant at level 0.05. This means that for these 3 variables, we can not reject the null hypotheses that the corrisponding coefficient is null. Therefore, we will abstain ourselves from interpreting those coefficients. We note that all the coefficients estimates $\beta_j$ have a positive sign, apart from strenght and marking. It seems normal that our continuous variables has a positive sign. Indeed, most of them are outlining player abilities, and it is clear that players who have better capabilities are on average better paid. Hwever, marking and strenght abilities seem to lower the salary, and we can not think of an explantion for that phenomenon. We can now take a look more precisely at our qualitative variables (Nationality and PreferredFoot). We see that the fact to be left-footed raise on average your salary by 0.02 thousand of dollars, i.e 20$. Since left-footed players compose only one fifth of all players, managers may want to keep them in the team, and therefore pay them more then the others. For the nationality variable, the Asian Nationality 
improve your salary by 280\$ (on average), the African Nationality by 100\$ and the Europeean Nationality by 120\$. We don't see any explication for that, especially for the $\beta_{Asian}$, since they don't have a so big football tradition.



# Coefficient Linear Combination

We will now test if a linear combination of our coeficients holds. We are interested in the effect of the overall variable. The objective will be to see if its effect is compensated by the effect of the other coefficients relative to player abilities. We will therefore test the following hypothesis : 
\[H_0 : \beta_{Overall} 
\\= \frac{\beta_{Potential}+\beta_{HeadingAccuracy}+\beta_{Reactions}+\beta_{Jumping}+\beta_{Strength}+
\beta_{Aggression}+\beta_{Composure}+\beta_{Marking}}{8} \]

```{r}
library("car")
hyp = c(0, -8, 1, 0, 1, 1, 1, 1 , 1, 1, 1, 0, 0, 0, 0, 0)
linearHypothesis(model, hypothesis.matrix = hyp)

```

We see in the table above that we can reject the restricted model (i.e the model under $H_0$). We can conlude that the effect of the overall variable is not the opposite effect of the mean of the effect of the players abilities. 




# Test of nullity of a subset of coefficients

We would like to see if we can simplify our model a bit to a restricted model. For that, we will check if we can say that some of the $\beta_j$ (actually, the ones that we have previously suspected to be non significative) are simultaneously equal to 0.
Our test staistic is 
\[\frac{(SSE_0 - SSE)/q}{SSE/(n-p)} \sim F_{q, n-p} \]
The idea behind this test is that under $H_0$, $SSE$ and $SSE_0$ should be close. We will therefore reject $H_0$ if the statistic is beyond a critical value (for a $F_{q, n-p}$ distribution).

```{r}
linearHypothesis(model, c("Jumping=0", "Weight=0", "Preferred_FootLeft=0"))
```

In the result table above, we look at the foolowing null hypothesis :
\[H_0 : \beta_{Jumping}=\beta_{Weight}=\beta_{PreferredFootLeft}=0\]

We see that we have a p-value of 0.04 for this F-test. That is, we reject the null hypothesis if we choose a confidence level of 0.05, but not if the level is 0.01. Since we have worked with a p-value of 0.05 since the begenning, we can reject $H_0$. So we reject the assertion that $\beta_{Jumping}$, $\beta_{Weight}$ and $\beta_{PreferredFootLeft}$ are null simultaneously.











